---
title: "R Notebook"
output: 
author: Julie Christine Freijat
---

**Data downloaded from:**

<https://data.kcmo.org/311/311-Reported-Issues-by-Neighborhood/nzmv-f45g> <https://data.kcmo.org/311/311-Call-Center-Service-Requests-2007-March-2021/7at3-sxhp> <https://data.kcmo.org/311/311-Call-Center-Reported-Issues/d4px-6rwg>

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

**Load in our data.**

We will be using 311 call center data from March 2021 to the July 28. In order to look at the call center data at a council district level, we need to export each council district separately, add an identifying column and then bind them together. Keep in mind - there is some data that exists that is not assigned to a council district. This data will be loaded in as NA. We also have a data set that looks at the incidents at a neighborhood level. We will also be using the historical 311 call center data that logs the cases from 2007 to 2021.

```{r}
district_1 <- read_csv("data/311_Call_Center_Reported_Issues_District_1.csv")
district_2 <- read_csv("data/311_Call_Center_Reported_Issues_District_2.csv")
district_3 <- read_csv("data/311_Call_Center_Reported_Issues_District_3.csv")
district_4 <- read_csv("data/311_Call_Center_Reported_Issues_District_4.csv")
district_5 <- read_csv("data/311_Call_Center_Reported_Issues_District_5.csv")
district_6 <- read_csv("data/311_Call_Center_Reported_Issues_District_6.csv")
district_na <- read_csv("data/311_Call_Center_Reported_Issues_District_NA.csv")

#Adding identifying columns
district_1 <- district_1 %>% add_column(district=1:1)
district_2 <- district_2 %>% add_column(district=2:2)
district_3 <- district_3 %>% add_column(district=3:3)
district_4 <- district_4 %>% add_column(district=4:4)
district_5 <- district_5 %>% add_column(district=5:5)
district_6 <- district_6 %>% add_column(district=6:6)
district_na <- district_na %>% add_column(district=0:0)

#Cleaning Names
district_1 <- clean_names(district_1)
district_2 <- clean_names(district_2)
district_3 <- clean_names(district_3)
district_4 <- clean_names(district_4)
district_5 <- clean_names(district_5)
district_6 <- clean_names(district_6)


#Binding the rows
all_districts <- bind_rows(district_1, district_2, district_3, district_4, district_5, district_6, district_na)

#Cleaning the names
all_districts <- clean_names(all_districts)

#Creating new date columns
all_districts <- all_districts %>% mutate(new_creation_date=mdy_hms(open_date_time))
all_districts <- all_districts %>% mutate(creation_year=(year(new_creation_date)))
all_districts <- all_districts %>% mutate(creation_month=(month(new_creation_date)))
all_districts <- all_districts %>% mutate(creation_day=(day(new_creation_date)))
all_districts <- all_districts %>% mutate(new_closed_date=mdy_hms(resolved_date))
all_districts <- all_districts %>% mutate(closed_year=(year(new_closed_date)))
all_districts <- all_districts %>% mutate(closed_month=(month(new_closed_date)))
all_districts <- all_districts %>% mutate(closed_day=(day(new_closed_date)))

#Doing a raw count to double check data 
all_districts %>% count()
raw_count_CD <- all_districts %>% group_by(district) %>% count()
```

```{r}
#Load in neighborhood data
nbhd <- read_csv("data/311nbhd-with-objectid.csv")
kc_nb <- read_csv("data/Kansas_City_Neighborhood_Borders.csv")

#Clean names and rename columns
nbhd <- clean_names(nbhd)
kc_nb <- clean_names(kc_nb)
kc_nb <- rename(kc_nb, neighborhood=nbhname)
kc_nb <- rename(kc_nb, nbhd_id=nbhid)

#Creating new date columns
nbhd <- nbhd %>% mutate(new_creation_date=mdy_hms(open_date_time))
nbhd <- nbhd %>% mutate(creation_year=(year(new_creation_date)))
nbhd <- nbhd %>% mutate(creation_month=(month(new_creation_date)))
nbhd <- nbhd %>% mutate(creation_day=(day(new_creation_date)))
nbhd <- nbhd %>% mutate(new_closed_date=mdy_hms(resolved_date))
nbhd <- nbhd %>% mutate(closed_year=(year(new_closed_date)))
nbhd <- nbhd %>% mutate(closed_month=(month(new_closed_date)))
nbhd <- nbhd %>% mutate(closed_day=(day(new_closed_date)))

#Doing a raw count of all cases by neighborhood
raw_count_nb <- nbhd %>% group_by(neighborhood, objectid) %>% count() %>% arrange(desc(n))
#Combining so I have other data if I want it
kc_nb_all <- right_join(raw_count_nb, kc_nb, by="objectid")
```

```{r}
#Load in historical data
kcmo_historical <- read_csv("data/311_Call_Center_Service_Requests__2007_-_March_2021.csv")

#Clean it
kcmo_historical <- clean_names(kcmo_historical)
```

**Calculating averages and medians for each council district.**

Some notes about how I calculated medians: The statuses of "canceled" and "referred" will not be included because they could indicate cases were not in the 311 jurisdiction, false or duplicates. I also excluded any days that are equal to NA or less than -1. When days open is equal to -1, it is due to a calculation error made when subtracting the open date from the resolved date. The open date records both the date and time of the incident, but the close date records only the date of when the incident is closed and sets the time equal to 12 a.m. For the purpose of calculations, anything greater than -1 and less than 1 were both counted as "0" days. Any cases with response times less than -1 were excluded (around 50 cases total).

Medians are used in the final analysis because the medians and averages are very different, which indicates the data has a skewed distribution. In this scenario, medians are a better representation of the data.

```{r}
cd_1 <- all_districts %>% filter(district == "1" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_1 <- cd_1 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_1 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 43.06115
cd_1 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=5

cd_2 <- all_districts %>% filter(district == "2" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_2 <- cd_2 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_2 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 43.39681
cd_2 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open)) 
#median=5

cd_3 <- all_districts %>% filter(district == "3" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_3 <- cd_3 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_3 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 61.3578
cd_3 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=7

cd_4 <- all_districts %>% filter(district == "4" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_4 <- cd_4 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_4 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 47.9106
cd_4 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=6

cd_5 <- all_districts %>% filter(district == "5" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_5 <- cd_5 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_5 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 55.405
cd_5 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=6

cd_6 <- all_districts %>% filter(district == "6" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
cd_6 <- cd_6 %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
cd_6 %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 38.17251
cd_6 %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=4


nbhd_median <- nbhd %>% filter(!current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
nbhd_median <- nbhd_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
nbhd_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))

```

**Raw incident count**

```{r}
all_districts %>% group_by(issue_type) %>% count %>% arrange(desc(n))
incident_raw_count <- all_districts %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
```

**Finding out the most common issues in different council districts**

```{r}
all_districts %>% filter(district=="1") %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
#trash/recycling/bulky services
#property violations
#street maintenance

all_districts %>% filter(district=="2") %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
#trash/recycling/bulky services
#street maintenance
#water service

all_districts %>% filter(district=="3") %>% group_by(issue_type) %>% count() %>% arrange(desc(n)) 
#property violations
#trash/recycling/bulky services
#illegal dumping

all_districts %>% filter(district=="4") %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
#property violations
#street maintenance
#trash/recycling/bulky services

all_districts %>% filter(district=="5") %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
#trash/recycling/bulky services
#property violations
#water service

all_districts %>% filter(district=="6") %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
#trash/recycling/bulky services
#street maintenance
#water service
```

```{r}
pv_median <- all_districts %>% filter(issue_type=="Property Violations" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
pv_median <- pv_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
pv_median %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 141.5752	
pv_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=55


#NOTE: To include all incidents in this category, we have to filter by two issues since one is misspelled. 
TRB_median <- all_districts %>% filter(issue_type=="Trash/Recycling and Bulky Services" | issue_type=="Trash / Recycling / Bulky Services" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
TRB_median <- TRB_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
TRB_median %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 63.55177	
TRB_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=2

water_median <- all_districts %>% filter(issue_type=="Water Service" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
water_median <- water_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
water_median %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 12.60814	
water_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=3

sm_median <- all_districts %>% filter(issue_type=="Street Maintenance" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
sm_median <- sm_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
sm_median %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open))
#average= 52.44108	
sm_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=21

id_median <- all_districts %>% filter(issue_type=="Illegal Dumping" & !current_status == "canceled" & !current_status== "referred" & days_open>=-1 & !is.na(days_open))
id_median <- id_median %>% mutate(new_days_open = replace(days_open, which(days_open<1), 0))
id_median %>% arrange(new_days_open) %>% summarise(average = mean(new_days_open)) 
#average= 18.15693	
id_median %>% select(new_days_open) %>% arrange(new_days_open) %>% summarise(total = median(new_days_open))
#median=7

all_districts %>% group_by(issue_type) %>% count() %>% arrange(desc(n))
```

**Making data frames for issues of interest**

```{r}
illegal_dumping <- all_districts %>% 
  group_by(district) %>% 
  filter(issue_type == "Illegal Dumping") %>% 
  count()

#NOTE: Again including both versions here. 
trash <- all_districts %>% 
  group_by(district) %>% 
  filter(issue_type == "Trash / Recycling / Bulky Services" | issue_type=="Trash/Recycling and Bulky Services") %>% 
  count()

streets <- all_districts %>% 
  group_by(district) %>% 
  filter(issue_type == "Street Maintenance") %>% 
  count()

property_violations <- all_districts %>% 
  group_by(district) %>% 
  filter(issue_type == "Property Violations") %>% 
  count()

water_service <- all_districts %>% 
  group_by(district) %>% 
  filter(issue_type == "Water Service") %>% 
  count()
```

**Aggregating historical data**

```{r}
kcmo_his_year <- kcmo_historical %>% select(creation_year)
kcmo_yr <- nbhd %>% select(creation_year)
kcmo_all_yrs <- bind_rows(kcmo_his_year, kcmo_yr)
kcmo_all_yrs <- kcmo_all_yrs %>% group_by(creation_year) %>% count()
```

**Creating a searchable database**

NOTE: We used object id's to combine the files because some of the neighborhoods are unnamed and can't be accurately combined except for with the object id.

```{r}
#Calculating median days open
median_days <- nbhd %>% 
  group_by(objectid) %>% 
  filter(!is.na(days_open) | days_open>-1)

median_days <- median_days %>% 
  mutate(new_days_open = replace(days_open, which(days_open<1), 0)) 

median_days <- median_days %>% 
  arrange(days_open) %>% 
  summarise(total = median(days_open))

median_days <- rename(median_days, median_days=total)

#Calculating number of reports each year 
count_over_time <- nbhd %>% 
  group_by(creation_year, objectid) %>% 
  count() %>% 
  pivot_wider(names_from=creation_year, values_from=n)

#Calculating number of reports each month each year
count_over_time_month <- nbhd %>% group_by(creation_year, creation_month, objectid) %>% count()
count_over_time_month <- count_over_time_month %>% 
  unite("creation_date", creation_month:creation_year, sep="-")
count_over_time_month <- count_over_time_month %>% 
  pivot_wider(names_from=creation_date, values_from=n)

#Join tables
kcmo_nb_data_table <- left_join(raw_count_nb, count_over_time, by="objectid")
kcmo_nb_data_table <- left_join(median_days, kcmo_nb_data_table, by="objectid")
kcmo_nb_data_table <- left_join(count_over_time_month, kcmo_nb_data_table, by="objectid")

#Counting issues  
issue_count_nb <- nbhd %>% 
  group_by(issue_type, objectid) %>% 
  count() %>% 
  arrange(desc(n))
issue_count_nb <- rename(issue_count_nb, value=n)
#Slicing top issue out
issue_count_nb <- issue_count_nb %>% 
  group_by(objectid) %>% 
  slice_max(value, n = 1)
issue_count_nb <- rename(issue_count_nb, issue_count = value)

#Joining top issue to table 
kcmo_nb_data_table <- left_join(issue_count_nb, kcmo_nb_data_table, by="objectid")

#The names for the unnamed neighborhoods are added into the visualization rather than in R.

```

**Writing CSV's for export.**

```{r}
write.csv(illegal_dumping, "data//illegal_dumping.csv", row.names=FALSE)
write.csv(trash, "data//trash.csv", row.names=FALSE)
write.csv(streets, "data//streets.csv", row.names=FALSE)
write.csv(property_violations, "data//property_violations.csv", row.names=FALSE)
write.csv(water_service, "data//water_service.csv", row.names=FALSE)
write.csv(raw_count_CD, "data//raw_count_CD.csv", row.names=FALSE)
write.csv(raw_count_nb, "data//raw_count_nb.csv", row.names=FALSE)
write.csv(kcmo_nb_data_table, "data//kcmo_nb_data_table.csv", row.names=FALSE)
write.csv(kcmo_all_yrs, "data//kcmo_all_yrs.csv", row.names=FALSE)
write.csv(incident_raw_count, "data//incident_raw_count.csv", row.names=FALSE)
write.csv(median_days, "data//median_days.csv", row.names=FALSE)

```
